Bridging the foundational gap: updating algebraic geometry in face of current challenges regarding formalizability, constructivity and predicativity

### 2021-11-21 slides-dagstuhl2021.tex
Slides finished(?)

>>>>

\begin{frame}{Foundational possibilities}
  \small

  \subhead{Schemes and scheme morphisms}
  \only<1-2>{
  \begin{itemize}
    \setlength\itemsep{-0.2em}
    %\item locally ringed spaces (or locales, distributive lattices, sites,
    %toposes, arithmetic universes), formal geometries
    \item locally ringed \ldots \\
    \ldots{} spaces \tabto{4cm}-- \textit{unconstructive, hard to formalize} \\
    \ldots{} locales \tabto{4cm}-- \textit{impredicative, superfluous opens} \\
    \ldots{} distributive lattices \tabto{4cm}-- \textit{need extension from basis} \\
    \ldots{} sites \tabto{4cm}-- \textit{current favorite! morphisms intricate} \\
    \ldots{} toposes \tabto{4cm}-- \textit{large structure} \\
    \ldots{} arithmetic universes \tabto{4cm}\mbox{-- \textit{a bit better;
    issue with relative spectrum}}
    \item formal geometries \tabto{4cm}-- \textit{can be regarded as sites}
    \item functor of points \tabto{4cm}-- \textit{large structure, issue with
    sizes or \\ \qquad
    with schemes not of finite presentation over the base ring}
    \item formal gluing data \tabto{4cm}-- \textit{morphisms intricate}
  \end{itemize}
  \textbf{A surprise of uncertain import:}
  \pause
  Internally to the big Zariski topos of a base scheme, the Zariski spectrum of
  a finitely presented algebra does have enough points! [Cherubini--Coquand]}
  \bigskip
  \pause

  \subhead{Zariski cohomology}
  \begin{itemize}
    \setlength\itemsep{-0.2em}
    \small
    \item Čech cohomology \tabto{4.5cm}-- \textit{ad hoc, but fine for}
    \tabto{4.5cm}\phantom{--} \textit{quasicompact separated schemes}
    \item injective resolutions \tabto{4.5cm}-- \textit{hopelessly unconstructive}
    \item flabby resolutions \tabto{4.5cm}-- \textit{probably unconstructive}
    \item pointwise Kan extensions \tabto{4.5cm}-- \textit{fine!}
  \end{itemize}
  \bigskip

  \subhead{Étale cohomology}
  \textcolor{red!90}{??}
\end{frame}
